---
title: '`r rstudioapi::getSourceEditorContext()$path %>% basename() %>% gsub(pattern = "\\.Rmd",replacement = "")`' 
author: "Avishai Wizel"
date: '`r Sys.time()`'
output: 
  html_notebook: 
    code_folding: hide
    toc: yes
    toc_collapse: yes
    toc_float: 
      collapsed: FALSE
    number_sections: true
    toc_depth: 1
---

## Parameters

```{r warning=FALSE}
suffix = ""
data_to_read = ""
```


## functions

```{r warning=FALSE}
source_from_github(repositoy = "DEG_functions",version = "0.2.24")
source_from_github(repositoy = "HMSC_functions",version = "0.1.12",script_name = "functions.R")
source_from_github(repositoy = "cNMF_functions",version = "0.3.72",script_name = "cnmf_function_Harmony.R")

no_neg <- function(x) {
  x = x + abs(min(x))
  x
}

sum_2_one <- function(x) {
  x =x/sum(x)
  x
}

```

## Data

```{r}

```



## Load object
```{python}
from cnmf import cNMF
import pickle
nfeatures = "2K"
f = open('./Data/cNMF/HMSC_cNMF_' + nfeatures+ 'vargenes/cnmf_obj.pckl', 'rb')
cnmf_obj = pickle.load(f)
f.close()
```


```{python}
selected_k = 4
density_threshold = 0.1
# cnmf_obj.consensus(k=selected_k, density_threshold=density_threshold,show_clustering=True)
usage_norm, gep_scores, gep_tpm, topgenes = cnmf_obj.load_results(K=selected_k, density_threshold=density_threshold)
```

```{r}
gep_scores = py$gep_scores
gep_tpm = py$gep_tpm
all_metagenes= py$usage_norm
```


```{r}
nfeatures = 2000
acc1_cancer_cells = FindVariableFeatures(object = acc1_cancer_cells,nfeatures = nfeatures)
vargenes = VariableFeatures(object = acc1_cancer_cells)
hmsc_expression = t(as.matrix(GetAssayData(acc1_cancer_cells,slot='data')))
hmsc_expression = 2**hmsc_expression #convert from log2(tpm+1) to tpm
hmsc_expression = hmsc_expression-1
hmsc_expression = hmsc_expression[,!colSums(hmsc_expression==0, na.rm=TRUE)==nrow(hmsc_expression)] #delete rows that have all 0
hmsc_expression = hmsc_expression[,vargenes]
hmsc_expression = hmsc_expression  %>% as.data.frame()


```






```{python}
def get_norm_counts(counts, tpm,high_variance_genes_filter): #from cnmf.py
    import numpy as np
    import scipy.sparse as sp
    ## Subset out high-variance genes
    norm_counts = counts[:, high_variance_genes_filter].copy()
  
    ## Scale genes to unit variance
    if sp.issparse(tpm.X):
        sc.pp.scale(norm_counts, zero_center=False)
        if np.isnan(norm_counts.X.data).sum() > 0:
            print('Warning NaNs in normalized counts matrix')                       
    else:
        norm_counts.X /= norm_counts.X.std(axis=0, ddof=1)
        if np.isnan(norm_counts.X).sum().sum() > 0:
            print('Warning NaNs in normalized counts matrix')                    
    
  
    ## Check for any cells that have 0 counts of the overdispersed genes
    zerocells = norm_counts.X.sum(axis=1)==0
    if zerocells.sum()>0:
        examples = norm_counts.obs.index[zerocells]
        print('Warning: %d cells have zero counts of overdispersed genes. E.g. %s' % (zerocells.sum(), examples[0]))
        print('Consensus step may not run when this is the case')
    
    return(norm_counts)
  
def get_usage_from_score(counts,tpm, genes,cnmf_obj):
      import anndata as ad
      import scanpy as sc
      import numpy as np
      from sklearn.decomposition import non_negative_factorization
      import pandas as pd
      counts_adata = ad.AnnData(counts)
      tpm_adata = ad.AnnData(tpm)
      norm_counts = get_norm_counts(counts=counts_adata,tpm=tpm_adata,high_variance_genes_filter=np.array(genes)) #norm counts as cnmf
      spectra = cnmf_obj.get_median_spectra(k=4) #get score 
      spectra = spectra[spectra.columns.intersection(genes)] #remove genes not in @genes
      usage_by_calc,_,_ = non_negative_factorization(X=norm_counts.X, H = spectra.values, update_H=False,n_components = 4,max_iter=1000,init ="random")
      usage_by_calc = pd.DataFrame(usage_by_calc, index=counts.index, columns=spectra.index) #insert to df+add names
      usage_by_calc = usage_by_calc.div(usage_by_calc.sum(axis=1), axis=0) # sum rows to 1 
      return(usage_by_calc)

```

```{python}

hmsc_expression = r.hmsc_expression
vargenes = r.vargenes
usage_by_calc = get_usage_from_score(counts=hmsc_expression,tpm=hmsc_expression,genes= vargenes,cnmf_obj=cnmf_obj)
```

```{r}
all_metagenes= py$usage_norm
all_metagenes_by_calc= py$usage_by_calc
all_metagenes_by_calc = all_metagenes_by_calc[,c(1,4,3,2)]
cor(all_metagenes,all_metagenes_by_calc)
```








```{r}
acc1_cancer_cells = FindVariableFeatures(object = acc1_cancer_cells,nfeatures = 2000)
vargenes = VariableFeatures(object = acc1_cancer_cells)

acc_expression = t(as.matrix(GetAssayData(acc_cancerCells_noACC1,slot='data')))
acc_expression = acc_expression[,vargenes]
acc_expression = 2**acc_expression #convert from log2(tpm+1) to tpm
acc_expression = acc_expression-1
# acc_expression = acc_expression[,!colSums(acc_expression==0, na.rm=TRUE)==nrow(acc_expression)] #delete rows that have all 0
acc_expression = acc_expression  %>% as.data.frame()

all_0_genes = colnames(acc_expression)[colSums(acc_expression==0, na.rm=TRUE)==nrow(acc_expression)] #delete rows that have all 0
vargenes = vargenes[!vargenes %in% all_0_genes]
```


```{python}
acc_expression = r.acc_expression
vargenes = r.vargenes
usage_by_calc = get_usage_from_score(counts=acc_expression,tpm=acc_expression,genes=np.array(vargenes),cnmf_obj=cnmf_obj)
```



# Hallmark Enrichment analysis by top 200 genes of each program
```{r fig.height=8, fig.width=8, results='hide'}
plt_list = list()
for (i in 1:ncol(gep_scores)) {
  top_genes = gep_scores  %>%  arrange(desc(gep_scores[i])) #sort by score a
  top = head(rownames(top_genes),200) #take top top_genes_num
  res = genes_vec_enrichment(genes = top,background = rownames(gep_scores),homer = T,title = 
                    i,silent = T,return_all = T)
   
  plt_list[[i]] = res$plt
}
gridExtra::grid.arrange(grobs = plt_list)
```

# Hallmark + canonical Enrichment analysis by top 200 genes of each program

```{r fig.height=8, fig.width=8, results='hide'}

canonical_pathways = msigdbr(species = "Homo sapiens", category = "C2") %>% dplyr::filter(gs_subcat != "CGP") %>%  dplyr::distinct(gs_name, gene_symbol) 

plt_list = list()
for (i in 1:ncol(gep_scores)) {
  top_genes = gep_scores  %>%  arrange(desc(gep_scores[i])) #sort by score a
  top = head(rownames(top_genes),200) #take top top_genes_num
  res = genes_vec_enrichment(genes = top,background = rownames(gep_scores),homer = T,title = 
                    i,silent = T,return_all = T,custom_pathways = canonical_pathways)
   
  plt_list[[i]] = res$plt
}
gridExtra::grid.arrange(grobs = plt_list)
```
# Expression by scores from cnmf {.tabset}
```{r fig.height=10, fig.width=10, results='asis'} 
all_metagenes = py$usage_by_calc
all_metagenes = scale(all_metagenes) %>% as.data.frame()
# Make metagene names
for (i in 1:ncol(all_metagenes)) {
  colnames(all_metagenes)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes)) {
  metage_metadata = all_metagenes %>% select(i)
  acc_cancerCells_noACC1 = AddMetaData(object = acc_cancerCells_noACC1,metadata = metage_metadata)
}

print_tab(plt = 
            FeaturePlot(object = acc_cancerCells_noACC1,features = colnames(all_metagenes))
,title = "expression",subtitle_num = 2)


print_tab(
  plt = FeaturePlot(object = acc_cancerCells_noACC1,features = colnames(all_metagenes),max.cutoff = 0.2)
          ,title = "max.cutoff = 0.2",subtitle_num = 2)

```
```{r}
FeaturePlot(object = acc_cancerCells_noACC1,features = "nFeature_RNA")
FeaturePlot(object = acc_cancerCells_noACC1,features = "MYB")
DimPlot(object = acc_cancerCells_noACC1,group.by = "orig.ident")
```




```{r}
acc_cancerCells_noACC1 = program_assignment(dataset = acc_cancerCells_noACC1,larger_by = 1.25,program_names = colnames(all_metagenes))
```



# UMAPS {.tabset}

```{r echo=TRUE, results='asis'}
colors =  rainbow(acc_cancerCells_noACC1$program.assignment %>% unique() %>% length()-1)
colors = c(colors,"grey")
print_tab(plt = DimPlot(object = acc_cancerCells_noACC1,group.by = "program.assignment",cols =colors),title = "program.assignment")
print_tab(plt = DimPlot(object = acc_cancerCells_noACC1,group.by = "patient.ident"),title = "patient.ident")

print_tab(plt = FeaturePlot(object = acc_cancerCells_noACC1,features = "lum_score"),title = "lum_score")
 

# Expression by scores from cnmf {.tabset}
```

```{r fig.height=10, fig.width=10, results='asis'} 
all_metagenes = py$usage_by_calc

# Make metagene names
for (i in 1:ncol(all_metagenes)) {
  colnames(all_metagenes)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes)) {
  metage_metadata = all_metagenes %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

print_tab(plt = 
            FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
,title = "expression",subtitle_num = 2)


print_tab(
  plt = FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes),max.cutoff = 0.05)
          ,title = "max.cutoff = 0.05",subtitle_num = 2)

```


# Expression by scores from cnmf + Z score {.tabset}
```{r fig.height=10, fig.width=10, results='asis'} 


all_metagenes_norm= scale(all_metagenes) %>% as.data.frame()

# Make metagene names
for (i in 1:ncol(all_metagenes_norm)) {
  colnames(all_metagenes_norm)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes_norm)) {
  metage_metadata = all_metagenes_norm %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

print_tab(plt = 
            FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
,title = "expression",subtitle_num = 2)


print_tab(
  plt = FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes),max.cutoff = 0.5)
          ,title = "max.cutoff = 0.5",subtitle_num = 2)

```
# Expression by coefficients inversion * expression  {.tabset}
```{r}

all_metagenes_inversed  = expression_inversion(gep_scores = gep_scores %>% as.matrix() ,dataset = acc1_cancer_cells) %>% t() %>% as.data.frame()
print(cor(all_metagenes_inversed,all_metagenes))
```

```{r fig.height=10, fig.width=10, results='asis'} 

# Make metagene names
for (i in 1:ncol(all_metagenes_inversed)) {
  colnames(all_metagenes_inversed)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes_inversed)) {
  metage_metadata = all_metagenes_inversed %>% select(i)
}

FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes_inversed))

```

# Expression by (coefficients inversion sum to 1) * expression {.tabset}
```{r}
gep_scores_norm = apply(gep_scores, 2, no_neg)
gep_scores_norm = apply(gep_scores_norm, 2, sum_2_one)


all_metagenes_inversed  = expression_inversion(gep_scores = gep_scores_norm ,dataset = acc1_cancer_cells) %>% t() %>% as.data.frame()
print(cor(all_metagenes_inversed,all_metagenes))

all_metagenes_inversed = apply(all_metagenes_inversed, 2, no_neg)
all_metagenes_inversed = apply(all_metagenes_inversed, 1, sum_2_one) %>% t() %>% as.data.frame()
print(cor(all_metagenes_inversed,all_metagenes))

```



```{r fig.height=10, fig.width=10, results='asis'} 

# Make metagene names
for (i in 1:ncol(all_metagenes_inversed)) {
  colnames(all_metagenes_inversed)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes_inversed)) {
  metage_metadata = all_metagenes_inversed %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

print_tab(plt = 
            FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
,title = "expression",subtitle_num = 2)


print_tab(
  plt = FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes),max.cutoff = 0.3)
          ,title = "max.cutoff = 750",subtitle_num = 2)
```

# Expression by coefficients sum to 1 * expression {.tabset}
```{r}
gep_scores_norm = apply(gep_scores, 2, no_neg)
gep_scores_norm = apply(gep_scores_norm, 2, sum_2_one) %>% as.data.frame()


all_metagenes_mult  = expression_mult(gep_scores = gep_scores_norm ,dataset = acc1_cancer_cells,top_genes = T)
print(cor(all_metagenes_mult,all_metagenes))

all_metagenes_mult = apply(all_metagenes_mult, 2, no_neg)
all_metagenes_mult = apply(all_metagenes_mult, 1, sum_2_one) %>% t() %>% as.data.frame()
print(cor(all_metagenes_mult,all_metagenes))
```

```{r fig.height=10, fig.width=10, results='asis'} 

# Make metagene names
for (i in 1:ncol(all_metagenes_mult)) {
  colnames(all_metagenes_mult)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes_mult)) {
  metage_metadata = all_metagenes_mult %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

print_tab(plt = 
            FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
,title = "expression",subtitle_num = 2)


print_tab(
  plt = FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes),max.cutoff = 50)
          ,title = "max.cutoff = 50",subtitle_num = 2)
```
