---
title: "R Notebook"
output: 
  html_notebook: 
    code_folding: hide
---
## Original heatmap
```{r fig.height=14, fig.width=10}
acc1_all_cells.df = acc1_all_cells@assays[["RNA"]]@data %>% as.data.frame()
acc1_cor = cor(acc1_all_cells.df)
annotation = FetchData(object = acc1_all_cells,vars = c("cell.type"))
pht1 = pheatmap(acc1_cor,annotation_row = annotation,fontsize = 6)
```
## Assign into clusters
```{r}
num_of_clusters = 3
clustering_distance = "euclidean"
myannotation = as.data.frame(cutree(pht1[["tree_row"]], k = num_of_clusters))
 
names(myannotation)[1] = "cluster"
  myannotation$cluster = as.factor(myannotation$cluster)
  
  palette1 <- palette(brewer.pal(num_of_clusters, "Paired"))
  palette1 <- palette(brewer.pal(num_of_clusters, "Paired")) #need to do twice to work
  
  names(palette1) = unique(myannotation$cluster)
  ann_colors = list (cluster = palette1)
  annotation = list(ann_colors = ann_colors, myannotation = myannotation)
  pheatmap(mat = acc1_cor,annotation_col =  annotation[["myannotation"]], annotation_colors = annotation[["ann_colors"]], clustering_distance_rows = clustering_distance,clustering_distance_cols = clustering_distance)
```
cluster 1+2 = normal
cluster 3 = outliers

## Normal vs outliers {.tabset}


### percent.mt

```{r}

#get cells from cluster 3:
outliers = myannotation %>% mutate(cluster = if_else(condition = cluster[row_number()]==3,true = "outliers",false = "normal"))
acc1_all_cells = AddMetaData(object = acc1_all_cells,metadata = outliers,col.name = "is_outlier")
comparisons <- list(c("outliers", "normal"))
VlnPlot(object = acc1_all_cells,group.by= "is_outlier",features = "percent.mt")+ stat_compare_means(comparisons = comparisons, label = "p.signif",label.y = 29)
```
### nCount_RNA
```{r}
#get cells from cluster 3:
outliers = myannotation %>% mutate(cluster = if_else(condition = cluster[row_number()]==3,true = "outliers",false = "normal"))
acc1_all_cells = AddMetaData(object = acc1_all_cells,metadata = outliers,col.name = "is_outlier")
comparisons <- list(c("outliers", "normal"))
VlnPlot(object = acc1_all_cells,group.by= "is_outlier",features = "nCount_RNA")+ stat_compare_means(comparisons = comparisons, label = "p.signif",label.y = 47000)
```
### nFeature_RNA
```{r}
#get cells from cluster 3:
outliers = myannotation %>% mutate(cluster = if_else(
  condition = cluster[row_number()]==3,
  true = "outliers",
  false = "normal"))
acc1_all_cells = AddMetaData(object = acc1_all_cells,metadata = outliers,col.name = "is_outlier")
comparisons <- list(c("outliers", "normal"))
VlnPlot(object = acc1_all_cells,group.by= "is_outlier",features = "nFeature_RNA")+ stat_compare_means(comparisons = comparisons, label = "p.signif",label.y = 13000)
```
## {-}

## heatmap with "correlation" distance

```{r fig.height=14, fig.width=10}
acc1_all_cells.df = acc1_all_cells@assays[["RNA"]]@data %>% as.data.frame()
acc1_cor = cor(acc1_all_cells.df)
annotation = FetchData(object = acc1_all_cells,vars = c("cell.type"))
pheatmap(acc1_cor,annotation_row = annotation,fontsize = 6,clustering_distance_cols = "correlation",clustering_distance_rows  = "correlation")
```


## correlation of all cells to myo markers found in article:
```{r}
myo_protein_markers = c("CNN1", "IL12B", "TP63","ACTA2")
expression = GetAssayData(object = acc1_cancer_cells,assay = "RNA",slot = "data") %>% as.data.frame()
myo_markers_expression = expression %>% dplyr::filter(row.names(expression) %in% myo_protein_markers)
expression = expression %>% t() %>% as.data.frame()
myo_markers_expression = myo_markers_expression %>% t() %>% as.data.frame()
```

```{r}
FeaturePlot(object = acc1_cancer_cells,features = myo_protein_markers,min.cutoff = 0)
```
IL12B is 0 in all cells

## top 10 correlated genes with TP63, ACTA2 and CNN1 combined
```{r}
myo_markers_expression = myo_markers_expression %>% select(-"IL12B") #remove IL12B since it has all 0
```

```{r results='hide'}
cor_mat = cor(expression, myo_markers_expression)
cor_mat = cor_mat[complete.cases(cor_mat), ] #remove rows with NA in at least one col
# pheatmap(cor_mat[,1,drop = F],cluster_rows = T,cluster_cols = F,show_rownames = F, show_colnames = T,treeheight_row = 0)
top_10_TP63 =   cor_mat %>% as.data.frame %>% select(1) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
top_10_ACTA2 =   cor_mat %>% as.data.frame %>% select(2) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
top_10_CNN1 =   cor_mat %>% as.data.frame %>% select(3) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
all_top_genes = c(top_10_TP63,top_10_ACTA2,top_10_CNN1)
```

Top correlated genes:

```{r}
all_top_genes
```

## enrichment analysis
```{r results='hide'}
genes_vec_enrichment(genes = all_top_genes,background = rownames(cor_mat),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "myo genes enrichment",add_msigdb_to_set = T)
```
## top 10 anti correlated genes with  TP63, ACTA2 and CNN1 combined and enrichment

```{r}
luminal_pathways = c("CHARAFE_BREAST_CANCER_LUMINAL_VS_BASAL_DN","CHARAFE_BREAST_CANCER_LUMINAL_VS_BASAL_UP","CHARAFE_BREAST_CANCER_LUMINAL_VS_MESENCHYMAL_DN","CHARAFE_BREAST_CANCER_LUMINAL_VS_MESENCHYMAL_UP","HUPER_BREAST_BASAL_VS_LUMINAL_DN","LIM_MAMMARY_LUMINAL_PROGENITOR_UP","SMID_BREAST_CANCER_LUMINAL_B_UP" )
# add luminal pathways
msigdb_gene_set = msigdbr(species = "Homo sapiens") 
msigdb_gene_set = msigdb_gene_set %>%as.data.frame() 
luminal_gs = msigdb_gene_set%>%  dplyr::filter(gs_name %in% luminal_pathways)
luminal_gs = luminal_gs %>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()

```

```{r warning=FALSE, results='hide'}
top_10_anticor_TP63 =   cor_mat %>% as.data.frame %>% select(1) %>%  arrange(desc(1)) %>% top_n(-10) %>% rownames()
top_10_anticor_ACTA2 =   cor_mat %>% as.data.frame %>% select(2) %>%  arrange(desc(1)) %>% top_n(-10) %>% rownames()
top_10_anticor_CNN1 =   cor_mat %>% as.data.frame %>% select(3) %>%  arrange(desc(1)) %>% top_n(-10) %>% rownames()
all_top_anticor_genes = c(top_10_anticor_TP63,top_10_anticor_ACTA2,top_10_anticor_CNN1)
genes_vec_enrichment(genes = all_top_anticor_genes,background = rownames(cor_mat),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "lum genes enrichment",add_msigdb_to_set = T,custom_pathways = luminal_gs)
```
```{r}
all_top_anticor_genes
```

## nFeature_RNA > 3000 filtring
after nFeature_RNA > 3000:
```{r}
acc1_all_cells.3000nFeature  <- subset(acc1_all_cells, subset = nFeature_RNA > 3000 )
comparisons <- list(c("outliers", "normal"))
VlnPlot(object = acc1_all_cells.3000nFeature,group.by= "is_outlier",features = "nFeature_RNA")+ stat_compare_means(comparisons = comparisons, label = "p.signif",label.y = 13000)
```
```{r}
print("cells in nFeature_RNA>2000:" %>% paste(ncol(acc1_all_cells)) )
print("cells in nFeature_RNA>3000:" %>% paste(ncol(acc1_all_cells.3000nFeature)) )

```

```{r}
#add outlier metadata to acc1 cancer cells:
outlier_metadata = FetchData(object = acc1_all_cells,vars = "is_outlier")
acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = outlier_metadata,col.name = "is_outlier")
acc1_cancer_cells.3000nFeature  <- subset(acc1_cancer_cells, subset = nFeature_RNA > 3000 )
```

```{r}
#get all genes expression and myo_protein_markers expression
myo_protein_markers = c("CNN1", "IL12B", "TP63","ACTA2")
expression = GetAssayData(object = acc1_cancer_cells.3000nFeature,assay = "RNA",slot = "data") %>% as.data.frame()
myo_markers_expression = expression %>% dplyr::filter(row.names(expression) %in% myo_protein_markers)
expression = expression %>% t() %>% as.data.frame()
myo_markers_expression = myo_markers_expression %>% t() %>% as.data.frame()
```

```{r}
myo_markers_expression = myo_markers_expression %>% select(-"IL12B") #remove IL12B since it has all 0
```

```{r results='hide'}
cor_mat = cor(expression, myo_markers_expression)
cor_mat = cor_mat[complete.cases(cor_mat), ] #remove rows with NA in at least one col
top_10_TP63 =   cor_mat %>% as.data.frame %>% select(1) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
top_10_ACTA2 =   cor_mat %>% as.data.frame %>% select(2) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
top_10_CNN1 =   cor_mat %>% as.data.frame %>% select(3) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
all_top_genes = c(top_10_TP63,top_10_ACTA2,top_10_CNN1)

```

```{r}
all_top_genes
```

```{r results='hide'}
genes_vec_enrichment(genes = all_top_genes,background = rownames(cor_mat),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "myo genes enrichment",add_msigdb_to_set = T)
```

```{r warning=FALSE, results='hide'}
top_10_anticor_TP63 =   cor_mat %>% as.data.frame %>% select(1) %>%  arrange(desc(1)) %>% top_n(-10) %>% rownames()
top_10_anticor_ACTA2 =   cor_mat %>% as.data.frame %>% select(2) %>%  arrange(desc(1)) %>% top_n(-10) %>% rownames()
top_10_anticor_CNN1 =   cor_mat %>% as.data.frame %>% select(3) %>%  arrange(desc(1)) %>% top_n(-10) %>% rownames()
all_top_anticor_genes = c(top_10_anticor_TP63,top_10_anticor_ACTA2,top_10_anticor_CNN1)
genes_vec_enrichment(genes = all_top_anticor_genes,background = rownames(cor_mat),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "lum genes enrichment",add_msigdb_to_set = T,custom_pathways = luminal_gs)
```




```{r}
all_top_anticor_genes
```

```{r fig.height=10, fig.width=10}
FeaturePlot(object = acc_cancer_cells,features = lum_genes)

```
```{r}
myoscore=apply(acc_cancer_cells@assays[["RNA"]][c("TP63","TP73","CDH3","KRT14","KRT5","ACTA2","CDH11","TAGLN","MYLK","DKK3","SPARC","TRIM29"),],2,mean)
lescore=apply(acc_cancer_cells@assays[["RNA"]][c("KIT","CLDN3","ANXA8","EHF","ELF5","KRT7","CLDN4","LCN2","LGALS3","WFDC2","ATP1B1","CD24"),],2,mean)
acc_cancer_cells=AddMetaData(acc_cancer_cells,lescore-myoscore,"luminal_over_myo")
```

```{r}
FeaturePlot(acc_cancer_cells,features = "luminal_over_myo")
```




