---
title: "Myo_lum_score"
author: "Avishai Wizel"
date: '`r Sys.time()`'
output: 
  html_notebook: 
    code_folding: hide
---
## Data

```{r}
acc1_cancer_cells = readRDS("./Data/" %>% paste0(data_to_read))
expression = GetAssayData(object = acc1_cancer_cells,assay = "RNA",slot = "data") %>% as.data.frame()

luminal_pathways = c("CHARAFE_BREAST_CANCER_LUMINAL_VS_BASAL_DN","CHARAFE_BREAST_CANCER_LUMINAL_VS_BASAL_UP","CHARAFE_BREAST_CANCER_LUMINAL_VS_MESENCHYMAL_DN","CHARAFE_BREAST_CANCER_LUMINAL_VS_MESENCHYMAL_UP","HUPER_BREAST_BASAL_VS_LUMINAL_DN","LIM_MAMMARY_LUMINAL_PROGENITOR_UP","SMID_BREAST_CANCER_LUMINAL_B_UP" )

# add luminal pathways
luminal_gs = msigdbr(species = "Homo sapiens") %>%as.data.frame() %>% dplyr::filter(gs_name %in% luminal_pathways)%>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()

```

## Parameters:

```{r}
suffix = "top_25_pearson_WithDistribution_withMyoAntiCor"
threshold = 25
data_to_read = "acc1_cancer_cells_15KnCount_V3.RDS"
```

-   suffix = `r suffix`
-   data_to_read = `r data_to_read`
-   threshold = `r threshold`

## functions

```{r warning=FALSE}
source("./.Rprofile")
source_from_github(repositoy = "DEG_functions",version = "0.2.1")

top_correlated <- function(genes, threshold,anti_cor = F) {
  markers_expression = FetchData(object = acc1_cancer_cells,vars = genes,slot = "data") #get genes expression
  markers_average = rowMeans(markers_expression) %>% as.data.frame() %>% rename("average" = 1) #average them
  cor_mat = cor(expression %>% t(), markers_average)%>% as.data.frame() #cor with all genes
  cor_mat = cor_mat[complete.cases(cor_mat),,drop=F]  %>% as.data.frame %>%  rename("corr" = 1) #remove rows with NA in at least one col
  if (threshold<1){ #if threshold is based on pearson correlation 
      if(anti_cor == T){top_genes =   cor_mat %>% as.data.frame %>% select(1) %>% dplyr::filter(.< threshold) %>% rownames()}else{
          top_genes =   cor_mat %>% as.data.frame %>% select(1) %>% dplyr::filter(.> threshold) %>% rownames()
      }
  }else{ #if threshold is based on top correlated genes 
      if(anti_cor == T){threshold  = threshold*(-1)}
      top_genes =   cor_mat %>%  top_n(threshold,corr) %>% rownames()
      }
  return(top_genes)
}

top_genes_cor_heatmap <- function(top_genes) {
  top_expression = expression %>% dplyr::filter(rownames(expression) %in% top_genes)
colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pht = pheatmap(mat = cor(top_expression %>% t(), top_expression %>% t()),color = my_palette, breaks = colors)
return(pht)
}

 
enriched_score_umap <- function(enrich_res, genes,col,distribution = F) {
  rownames(enrich_res) = enrich_res$pathway_name
  enriched_genes = enrich_res[col,"geneID"] %>% strsplit(split = "/") %>% .[[1]] %>% c(.,genes) #add original markers
  enriched_genes_score=apply(acc1_cancer_cells@assays$RNA@data[enriched_genes,],2,mean)
  acc1_cancer_cells=AddMetaData(acc1_cancer_cells,enriched_genes_score,"enriched_genes_score")
  if (distribution == F) {
      print(FeaturePlot(object = acc1_cancer_cells,features = "enriched_genes_score"))
    }else{
        data = FetchData(object = acc1_cancer_cells,vars = "enriched_genes_score")
        print(
          data %>% 
          ggplot(aes( x=enriched_genes_score)) + 
          geom_density() 
        )
    }
    return(list( score = enriched_genes_score,genes = enriched_genes))
}

```




# **1.Myo markers**

## 1.a) expression

```{r}
myo_protein_markers = c("CNN1", "IL12B", "TP63","ACTA2")
FeaturePlot(object = acc1_cancer_cells,features = myo_protein_markers,min.cutoff = 0)
```

IL12B is 0 in all cells
```{r}
myo_protein_markers = c("CNN1", "TP63","ACTA2")
```

## 1.b) top `r threshold` correlated genes with TP63+ACTA2+CNN1 {.tabset}

```{r}
top_myo  = top_correlated(genes = myo_protein_markers,threshold = threshold)
```

### Top correlated genes

```{r}
print("Number of genes = " %>% paste(length(top_myo)))
top_myo %>% head(30)
```


### Correlation

```{r fig.height=6, fig.width=8}
top_genes_cor_heatmap(top_genes = top_myo)
```

### Enrichment analysis

```{r results='hide'}
enrich_res = genes_vec_enrichment(genes = top_myo,background = rownames(acc1_cancer_cells),homer = T,title = "myo top enrichment")
enrich_res
```

### UMAP score of enriched genes

```{r}
lst = enriched_score_umap(enrich_res = enrich_res,genes = myo_protein_markers,col = 1)
myo_score = lst[["score"]]
enriched_genes = lst[["genes"]]
```
### Distribution of score

```{r}
lst = enriched_score_umap(enrich_res = enrich_res,genes = myo_protein_markers,col = 1,distribution = T)
```
genes in score:
```{r}
lst[["genes"]]
```
[CALD1](https://journals.sagepub.com/doi/epdf/10.1177/106689690501300108)
[Lysyl oxidase gene (LOXL2) expression in the stromal reaction to in situ and invasive ductal breast carcinoma](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1858268/)
TPM2 = tropomyosin 
##  {.unnumbered}


## Anti correlation to myo genes {.tabset}

### KIT anti correlation value
In all genes score:
```{r}
markers_expression = FetchData(object = acc1_cancer_cells,vars = top_myo,slot = "data") #get genes expression
markers_average = rowMeans(markers_expression) %>% as.data.frame() %>% rename("average" = 1) #average them
cor_mat = cor(expression %>% t(), markers_average)%>% as.data.frame() #cor with all genes
cor_mat = cor_mat[complete.cases(cor_mat),,drop=F]  %>% as.data.frame %>%  rename("corr" = 1) #remove rows with NA in at least one col
cor_mat = cor_mat %>% arrange(.[[1]]) 
kit_location = which(rownames(cor_mat) == "KIT")
kit_cor = cor_mat[rownames(cor_mat) == "KIT",1]
print(paste("location of KIT:", kit_location))
print(paste("correlation of KIT:", kit_cor))

```
In enriched genes:
```{r}
markers_expression = FetchData(object = acc1_cancer_cells,vars = enriched_genes,slot = "data") #get genes expression
markers_average = rowMeans(markers_expression) %>% as.data.frame() %>% rename("average" = 1) #average them
cor_mat = cor(expression %>% t(), markers_average)%>% as.data.frame() #cor with all genes
cor_mat = cor_mat[complete.cases(cor_mat),,drop=F]  %>% as.data.frame %>%  rename("corr" = 1) #remove rows with NA in at least one col
cor_mat = cor_mat %>% arrange(.[[1]]) 
kit_location = which(rownames(cor_mat) == "KIT")
kit_cor = cor_mat[rownames(cor_mat) == "KIT",1]
print(paste("location of KIT:", kit_location))
print(paste("correlation of KIT:", kit_cor))

```


### top 25 anti cor to myo score
```{r}
lum_threshold = 25
anti_cor  = top_correlated(genes = top_myo,threshold = lum_threshold,anti_cor = T)
anti_cor
if("KIT" %in% anti_cor){
  message("KIT is in genes list")
}else{message("KIT is NOT genes list")}
```

```{r results='hide'}
enrich_res = genes_vec_enrichment(genes = anti_cor,background = rownames(acc1_cancer_cells),homer = T,title = "lum top enrichment",custom_pathways = luminal_gs)
enrich_res
```

### top 50 anti cor to myo score

```{r}
lum_threshold = 50
anti_cor  = top_correlated(genes = top_myo,threshold = lum_threshold,anti_cor = T)
anti_cor
if("KIT" %in% anti_cor){
  message("KIT is in genes list")
}else{message("KIT is NOT genes list")}
```

```{r results='hide'}
enrich_res = genes_vec_enrichment(genes = anti_cor,background = rownames(acc1_cancer_cells),homer = T,title = "lum top enrichment",custom_pathways = luminal_gs)
enrich_res
```

### top 25 anti cor to myo enriched genes
```{r}
lum_threshold = 25
anti_cor  = top_correlated(genes =  enriched_genes,threshold = lum_threshold,anti_cor = T)
anti_cor
if("KIT" %in% anti_cor){
  message("KIT is in genes list")
}else{message("KIT is NOT genes list")}
```

```{r results='hide'}
enrich_res = genes_vec_enrichment(genes = anti_cor,background = rownames(acc1_cancer_cells),homer = T,title = "lum top enrichment",custom_pathways = luminal_gs)
enrich_res
```

### top 50 anti cor to myo enriched genes

```{r}
lum_threshold = 50
anti_cor  = top_correlated(genes = enriched_genes,threshold = lum_threshold,anti_cor = T)
anti_cor
if("KIT" %in% anti_cor){
  message("KIT is in genes list")
}else{message("KIT is NOT genes list")}
```

```{r results='hide'}
enrich_res = genes_vec_enrichment(genes = anti_cor,background = rownames(acc1_cancer_cells),homer = T,title = "lum top enrichment",custom_pathways = luminal_gs)
enrich_res
```

## {-}


# **2. Luminal markers**

## 2.a) expression

```{r}
lum_protein_markers = c("KIT")
FeaturePlot(object = acc1_cancer_cells,features = lum_protein_markers,min.cutoff = 0)
```

## 2 top  correlated genes with KIT {.tabset}

```{r}
top_kit = top_correlated(genes = lum_protein_markers,threshold = threshold)
```

### Top correlated genes

```{r}
print("#genes = " %>% paste(length(top_kit)))
top_kit
```

### Correlation

```{r fig.height=6, fig.width=8}
top_genes_cor_heatmap(top_genes = top_kit)
```

### Enrichment analysis

```{r results='hide'}
enrich_res = genes_vec_enrichment(genes = top_kit,background = rownames(expression),homer = T,title = "lum top enrichment",custom_pathways = luminal_gs)
enrich_res
```

### UMAP score of enriched genes

```{r}
lst = enriched_score_umap(enrich_res = enrich_res,genes = lum_protein_markers,col = 1)
lum_score = lst[["score"]]
enriched_genes = lst[["genes"]]
```
### Distribution of score

```{r}
lst = enriched_score_umap(enrich_res = enrich_res,genes = lum_protein_markers,col = 1,distribution = T)
```
genes in score:
```{r}
lst[["genes"]]
```
[CCND1 Copy Number Variation in Circulating Tumor DNA from Luminal B Breast Cancer Patients](https://pubmed.ncbi.nlm.nih.gov/35896251/)

##  {.unnumbered}

## Correlation between luminal and myo scores
```{r}
cor(lum_score,myo_score)
```



