---
title: "R Notebook"
output: html_notebook
---

```{r}
source_from_github(repositoy = "DEG_functions",version = "0.1.10")
acc_cancercells_original = read_rds(file = "./Data/acc_cancer_no146_primaryonly15k_cancercells.rds")
```

```{r fig.height=7, fig.width=7}
DimPlot(acc_cancercells_original,group.by  = "patient.ident")
```


```{r}
myo_genes =  c("TP63","TP73","CDH3","KRT14","KRT5","ACTA2","CDH11","TAGLN","MYLK","DKK3","SPARC","TRIM29")
lum_genes = c("KIT","CLDN3","ANXA8","EHF","ELF5","KRT7","CLDN4","LCN2","LGALS3","WFDC2","ATP1B1","CD24")
```

## Lum/ myo markers in original ACC {.tabset}




### Luminal genes:

```{r}
lum_markers  = FetchData(object = acc_cancercells_original, vars = lum_genes)
lum_cor = cor(lum_markers)

colors <- c(seq(-1,1,by=0.01))

my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")

pheatmap(lum_cor,clustering_distance_cols = "correlation", clustering_distance_rows = "correlation", color = my_palette, breaks = colors,display_numbers = T)

```

### Myo genes:


```{r}
myo_markers  = FetchData(object = acc_cancercells_original, vars = myo_genes)
myo_cor = cor(myo_markers)
pheatmap(myo_cor)

```
## {-}


## with ACC1 {.tabset}




### Luminal genes:

```{r}
lum_markers  = FetchData(object = acc1_cancer_cells, vars = lum_genes)
lum_cor = cor(lum_markers)


colors <- c(seq(-1,1,by=0.01))

my_palette <- c("green",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")

pheatmap(lum_cor,clustering_distance_cols = "correlation", clustering_distance_rows = "correlation", color = my_palette, breaks = colors,display_numbers = T)
```

### Myo genes:


```{r}
myo_markers  = FetchData(object = acc1_cancer_cells, vars = myo_genes)
myo_cor = cor(myo_markers)
pheatmap(myo_cor)

```

## {-}

```{r}
ElbowPlot(acc1_cancer_cells)

```

```{r}
acc1_cancer_cells = subset(x = acc_cancer_cells,subset = patient.ident == "ACC1")
acc1_cancer_cells <- FindVariableFeatures(acc1_cancer_cells, selection.method = "vst", nfeatures = 15000)
acc1_cancer_cells <- ScaleData(acc1_cancer_cells, vars.to.regress = c("percent.mt","nCount_RNA"))
acc1_cancer_cells <- RunPCA(acc1_cancer_cells, features = VariableFeatures(object = acc_cancer_cells))
acc1_cancer_cells <- FindNeighbors(acc1_cancer_cells, dims = 1:20,n.trees = 5)
acc1_cancer_cells <- FindClusters(acc1_cancer_cells,  resolution = 0.5)
```




```{r}
acc1_cancer_cells <- RunUMAP(acc1_cancer_cells,n.neighbors = 5,graph = 'RNA_snn')
DimPlot(acc1_cancer_cells)
```
```{r}

feature_to_test = filter_features(object = acc1_cancer_cells,ident.1 = 0,ident.2 = c(1,2,3))
markers = FindAllMarkers(object = acc1_cancer_cells,densify = T)
```

```{r}
test = FindMarkers(object = acc1_cancer_cells,ident.1 = 0,ident.2 = c(1,2,3),features =  feature_to_test,densify = T,logfc.threshold = 0)
```

```{r}
test_sig = test %>% dplyr::filter(p_val_adj<0.05) %>% rownames()
  genes_vec_enrichment(test_sig,gene_sets = "homer_hallmark", title = cluster_num, convert_background = T, add_msigdb_to_set = T,background = feature_to_test)
```

```{r}

markers_lst = list()
for (cluster_num in 0:3) {
feature_to_test = filter_features(object = acc1_cancer_cells,ident.1 = cluster_num,ident.2 = NULL)
test = FindMarkers(object = acc1_cancer_cells,ident.1 = cluster_num,features =  feature_to_test,densify = T,logfc.threshold = 0)
markers_lst[[cluster_num %>% as.character()]] = test

}

for (lst in markers_lst) {
test_sig = lst %>% dplyr::filter(p_val<0.05) %>% rownames()
  genes_vec_enrichment(test_sig,gene_sets = "homer_hallmark", title = cluster_num, convert_background = T, add_msigdb_to_set = T,background = feature_to_test)
}

```

```{r}
myoscore=apply(acc1_cancer_cells@assays[["RNA"]][c("TP63","TP73","CDH3","KRT14","KRT5","ACTA2","CDH11","TAGLN","MYLK","DKK3","SPARC","TRIM29"),],2,mean)
lescore=apply(acc1_cancer_cells@assays[["RNA"]][c("KIT","CLDN3","ANXA8","EHF","ELF5","KRT7","CLDN4","LCN2","LGALS3","WFDC2","ATP1B1","CD24"),],2,mean)
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,lescore-myoscore,"luminal_over_myo")
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,lescore,"luminal_score")
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,myoscore,"myo_score")

```

```{r}
FeaturePlot(acc1_cancer_cells,features = "luminal_over_myo")
```


```{r}
FeaturePlot(acc1_cancer_cells,features = "luminal_score",min.cutoff = 0)
FeaturePlot(acc1_cancer_cells,features = "myo_score",min.cutoff = 0)
```



```{r fig.height=7, fig.width=9}
FeaturePlot(object = acc1_cancer_cells,features = lum_genes)
```
```{r fig.height=7, fig.width=9}
FeaturePlot(object = acc1_cancer_cells,features = myo_genes)
```

```{r}
myoscore=apply(acc1_cancer_cells@assays[["RNA"]][c("TP63","TP73","CDH3","ACTA2","CDH11","TAGLN","MYLK","DKK3"),],2,mean)
lescore=apply(acc1_cancer_cells@assays[["RNA"]][c("KIT","CLDN3","ANXA8","ELF5","LCN2","WFDC2","ATP1B1"),],2,mean)
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,lescore-myoscore,"luminal_over_myo")
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,lescore,"luminal_score")
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,myoscore,"myo_score")
```

```{r}
FeaturePlot(acc1_cancer_cells,features = "luminal_score",min.cutoff = 0)
FeaturePlot(acc1_cancer_cells,features = "myo_score",min.cutoff = 0)
FeaturePlot(acc1_cancer_cells,features = "luminal_over_myo",cols = c("blue","grey","yellow"))
```

Options for markers:
1. hand picked like here above
2. most corr with CNN1 etc (need to check that with new UMAP)
3. maybe try to improve umap/ umap clustering

estimate n:
```{r}
estim.r <- nmf(acc1_data, 2:6, nrun=1, seed=123456,.options='v')

plot(estim.r)
consensusmap(estim.r, annCol=esGolub, labCol=NA, labRow=NA)
```

```{r}
library(NMF)
acc1_cancer_cells = FindVariableFeatures(object = acc1_cancer_cells,nfeatures = 10000)
acc1_data = GetAssayData(object = acc1_cancer_cells,assay = "RNA",slot = "data") %>% as.data.frame()
acc1_data = acc1_data[rownames(acc1_data) %in% VariableFeatures(object = acc1_cancer_cells), ]
  
acc1_data = acc1_data[rowSums(acc1_data[])>0,] #filter rows with all 0 (required for nmf)
res <- nmf(acc1_data, 4)
```


```{r}
w2 <- basis(res) %>% as.data.frame()
h2 <- coef(res)%>% as.data.frame()

```


## enrichment of metagene based on top X genes selected by "max" method

```{r}
top_features = extractFeatures(object = res,method = "max")
for (i in 1:length(top_features)) {
  top_features[[i]]= rownames(res)[top_features[[i]]]
}
top_features
```

```{r}
for (i in seq_along(top_features)) {
  if(i==3){next}
  gene_list = top_features[[i]]
  genes_vec_enrichment(genes = gene_list,background = rownames(res),gene_sets = "homer_hallmark",title = i,convert_background = T,add_msigdb_to_set = T,custom_pathways = luminal_gs)
}
```
## add score for each cell

```{r}
all_metagenes = h %>% t() %>% as.data.frame()
for (i in 1:ncol(all_metagenes)) {
  colnames(all_metagenes)[i] = "metagene." %>% paste0(i)
}

for (i in 1:ncol(all_metagenes)) {
  metage_metadata = all_metagenes %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
```


## same with n = 6

```{r}
library(NMF)
res_n6 <- nmf(acc1_data, 6)
```


```{r}
w_6 <- basis(res_n6) %>% as.data.frame()
h_6 <- coef(res_n6)%>% as.data.frame()
```


## enrichment of metagene based on top X genes selected by "max" method

```{r}
top_features = extractFeatures(object = res_n6,method = "max")
for (i in 1:length(top_features)) {
  top_features[[i]]= rownames(res_n6)[top_features[[i]]]
}
top_features
```


## enrichment of metagene based on top X genes
```{r}
num = 2
metagene1 <- w[order(-w[,num]),]
geneList = metagene1 %>% pull(num)
names(geneList) = rownames(metagene1)
m_t2g <- msigdbr(species = "Homo sapiens", category = "H") %>% 
  dplyr::select(gs_name, gene_symbol)

gsea_res = GSEA(geneList = geneList,TERM2GENE = m_t2g)
```

```{r}
num = 4
metagene1 <- w[order(-w[,num]),]
geneList = metagene1 %>% rownames()
top_list = geneList[1:20]
genes_vec_enrichment(genes = top_list,background = geneList,gene_sets = "homer_hallmark",title = "test",convert_background = T,add_msigdb_to_set = T,custom_pathways = luminal_gs)
```
CNMF:
```{r}
acc.data_for_cnmf = readRDS(file = "./Data/cnmf/acc.data_for_cnmf.rds")
acc.tpm_for_cnmf = readRDS(file = "./Data/cnmf/acc.tpm_for_cnmf.rds")

acc.data_for_cnmf = CreateSeuratObject(counts = acc.data_for_cnmf,min.cells = 0,min.features = 0) #convert to seurat object to modify gene names to be like original data
acc.data_for_cnmf = GetAssayData(object = acc.data_for_cnmf,assay = "RNA", slot = "data") #return to df
acc.data_for_cnmf = acc.data_for_cnmf[rownames(acc.data_for_cnmf) %in% rownames(acc1_cancer_cells), colnames(acc.data_for_cnmf) %in% colnames(acc1_cancer_cells)] %>% as.data.frame() #filter genes and cells like in the original dataset

acc.tpm_for_cnmf = CreateSeuratObject(counts = acc.tpm_for_cnmf,min.cells = 0,min.features = 0) #convert to seurat object to modify gene names in cnmf data
acc.tpm_for_cnmf = GetAssayData(object = acc.tpm_for_cnmf,assay = "RNA", slot = "data") #return to df
acc.tpm_for_cnmf = acc.tpm_for_cnmf[rownames(acc.tpm_for_cnmf) %in% rownames(acc1_cancer_cells), colnames(acc.tpm_for_cnmf) %in% colnames(acc1_cancer_cells)] %>% as.data.frame() #filter genes and cells like in the original dataset

#remove genes with all 0 (cnmf requirement)
zero_genes = rownames(acc.data_for_cnmf)[rowSums(acc.data_for_cnmf) == 0 ]
acc.tpm_for_cnmf = acc.tpm_for_cnmf[!rownames(acc.tpm_for_cnmf) %in% zero_genes, ]
acc.data_for_cnmf = acc.data_for_cnmf[!rownames(acc.data_for_cnmf) %in% zero_genes, ]

acc.tpm_for_cnmf = as.data.frame(t(acc.tpm_for_cnmf)) #cNMF require cellsXgenes 
write.table(x = acc.tpm_for_cnmf, file = "./Data/cnmf_input/acc_tpm_final.txt",sep = '\t',row.names = T,col.names=NA)

acc.data_for_cnmf = as.data.frame(t(acc.data_for_cnmf)) #cNMF require cellsXgenes 
write.table(x = acc.data_for_cnmf, file = "./Data/cnmf_input/acc_counts_final.txt",sep = '\t',row.names = T,col.names=NA)


```

```{python}
import cnmf
from cnmf import cNMF
import os
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
from IPython.display import Image
from matplotlib import gridspec
from importlib import reload


numiter=200 ## Set this to a larger value (200) for real data. We set this to a low value here for illustration
numworkers=1 ## Set this to a larger value and use the parallel code cells to try out parallelization
# numhvgenes=1500 ## Number of over-dispersed genes to use for running the factorizations
numhvgenes=2000 ## Number of over-dispersed genes to use for running the factorizations

K  = np.arange(3,5)
K  = 4

output_directory = './Data/cNMF'
run_name = 'try1'


countfn ='./Data/cnmf_input/acc_counts_final.txt'
tpmfn = './Data/cnmf_input/acc_tpm_final.txt'

cnmf_obj = cNMF(output_dir=output_directory, name=run_name)
cnmf_obj.prepare(counts_fn=countfn, components=K, n_iter=numiter, num_highvar_genes=numhvgenes,tpm_fn=tpmfn)

cnmf_obj.factorize(worker_i=0, total_workers=1)
cnmf_obj.combine()
cnmf_obj.k_selection_plot()



selected_K = 4
cnmf_obj.consensus(k=selected_K, density_threshold=0.1, show_clustering=True)
  
```

```{r}
h_cnmf = fread(input = "./Data/cNMF/try1/try1.gene_spectra_score.k_4.dt_0_1.txt",sep = "\t",sep2 = "\n",drop = 1)
```

```{r}
all_metagenes = fread(input = "./Data/cNMF/try1/try1.usages.k_4.dt_0_1.consensus.txt",sep = "\t",sep2 = "\n",header = T) %>% as.data.frame()
row.names(all_metagenes) = all_metagenes %>% pull(1)
all_metagenes[1]<- NULL

for (i in 1:ncol(all_metagenes)) {
  colnames(all_metagenes)[i] = "metagene." %>% paste0(i)
}

for (i in 1:ncol(all_metagenes)) {
  metage_metadata = all_metagenes %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
```
```{python}
import cnmf
from cnmf import cNMF
import os
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
from IPython.display import Image
from matplotlib import gridspec
from importlib import reload


numiter=200 ## Set this to a larger value (200) for real data. We set this to a low value here for illustration
numworkers=2 ## Set this to a larger value and use the parallel code cells to try out parallelization
# numhvgenes=1500 ## Number of over-dispersed genes to use for running the factorizations
numhvgenes=2000 ## Number of over-dispersed genes to use for running the factorizations

K  = np.arange(3,5)
K  = 4

output_directory = './Data/cNMF'
run_name = 'try2'


countfn ='./Data/cnmf_input/acc_tpm_final.txt'
tpmfn = './Data/cnmf_input/acc_tpm_final.txt'

cnmf_obj = cNMF(output_dir=output_directory, name=run_name)
cnmf_obj.prepare(counts_fn=countfn, components=K, n_iter=numiter, num_highvar_genes=numhvgenes,tpm_fn=tpmfn)

cnmf_obj.factorize(worker_i=0, total_workers=1)
cnmf_obj.combine()
cnmf_obj.k_selection_plot()



selected_K = 4
cnmf_obj.consensus(k=selected_K, density_threshold=0.1, show_clustering=True)
usage_norm, gep_scores, gep_tpm, topgenes = cnmf_obj.load_results(K=selected_K, density_threshold=0.1)
usage_norm.to_csv('./Data/cNMF/try2/usage_norm.csv')
```



```{r}
all_metagenes = fread(input = "./Data/cNMF/try2/usage_norm.csv",sep = ",",sep2 = "\n",header = T) %>% as.data.frame()

#add rownames
row.names(all_metagenes) = all_metagenes %>% pull(1)
all_metagenes[1]<- NULL

#Make metagene names
for (i in 1:ncol(all_metagenes)) {
  colnames(all_metagenes)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes)) {
  metage_metadata = all_metagenes %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
```

Try3- filter genes that detected in small number of cells

```{r}
acc.data_for_cnmf = read.table(file ="./Data/cnmf_input/acc_counts_final.txt",sep = "\t" ,header = T,row.names = 1)
zero_genes  = acc.data_for_cnmf %>% summarise_each(funs(sum(.==0))) #count 0
(zero_genes[1,] >= 141) %>% table() #from total of 144 cells
genes_to_omit =colnames(zero_genes)[zero_genes[1,] >= 141]

#load TPM and filter 
acc.tpm_for_cnmf = read.table(file = "./Data/cnmf_input/acc_tpm_final.txt",sep = '\t',row.names = 1,header = T)
acc.tpm_for_cnmf = acc.tpm_for_cnmf[,!colnames(acc.tpm_for_cnmf) %in% genes_to_omit]
write.table(x = acc.tpm_for_cnmf, file = "./Data/cnmf_input/acc_tpm_final_gene_filtered.txt",sep = '\t',row.names = T,col.names=NA)

```


```{python}
import cnmf
from cnmf import cNMF
import os
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
from IPython.display import Image
from matplotlib import gridspec
from importlib import reload


numiter=200 ## Set this to a larger value (200) for real data. We set this to a low value here for illustration
numworkers=1 ## Set this to a larger value and use the parallel code cells to try out parallelization
# numhvgenes=1500 ## Number of over-dispersed genes to use for running the factorizations
numhvgenes=2000 ## Number of over-dispersed genes to use for running the factorizations

K  = np.arange(3,10)
K  = 4

output_directory = './Data/cNMF'
run_name = 'try3'


countfn ='./Data/cnmf_input/acc_tpm_final_gene_filtered.txt'
tpmfn = './Data/cnmf_input/acc_tpm_final_gene_filtered.txt'

cnmf_obj = cNMF(output_dir=output_directory, name=run_name)
cnmf_obj.prepare(counts_fn=countfn, components=K, n_iter=numiter, num_highvar_genes=numhvgenes,tpm_fn=tpmfn)

cnmf_obj.factorize(worker_i=0, total_workers=1)
cnmf_obj.combine()
cnmf_obj.k_selection_plot()



selected_K = 6
cnmf_obj.consensus(k=selected_K, density_threshold=0.1, show_clustering=True)
usage_norm, gep_scores, gep_tpm, topgenes = cnmf_obj.load_results(K=selected_K, density_threshold=0.1)
usage_norm.to_csv('./Data/cNMF/try3/usage_norm.csv')
```



```{r}
all_metagenes = fread(input = "./Data/cNMF/try3/usage_norm.csv",sep = ",",sep2 = "\n",header = T,) %>% as.data.frame()

#add rownames
row.names(all_metagenes) = all_metagenes %>% pull(1)
all_metagenes[1]<- NULL

#Make metagene names
for (i in 1:ncol(all_metagenes)) {
  colnames(all_metagenes)[i] = "metagene." %>% paste0(i)
}

#add each metagene to metadata
for (i in 1:ncol(all_metagenes)) {
  metage_metadata = all_metagenes %>% select(i)
  acc1_cancer_cells = AddMetaData(object = acc1_cancer_cells,metadata = metage_metadata)
}

FeaturePlot(object = acc1_cancer_cells,features = colnames(all_metagenes))
```