---
title: "R Notebook"
output: 
  html_notebook: 
    code_folding: hide
---
```{r echo=TRUE,class.source = 'fold-show'}
source_from_github(repositoy = "DEG_functions",version = "0.1.10")
acc1_cancer_cells <- RunUMAP(acc1_cancer_cells, dims = 1:20,n.neighbors = 5)
DimPlot(object = acc1_cancer_cells)
```

# **Option 1- most correlated with cell markers from Bishop et al:**

# **1.Myo markers**
## 1.a) expression
```{r}
myo_protein_markers = c("CNN1", "IL12B", "TP63","ACTA2")
expression = GetAssayData(object = acc1_cancer_cells,assay = "RNA",slot = "data") %>% as.data.frame()
myo_markers_expression = expression %>% dplyr::filter(row.names(expression) %in% myo_protein_markers) %>% as.data.frame()
```

```{r}
FeaturePlot(object = acc1_cancer_cells,features = myo_protein_markers,min.cutoff = 0)
```

IL12B is 0 in all cells

## 1.b) top 10 correlated genes with TP63, ACTA2 and CNN1 combined {.tabset}

```{r}
myo_markers_expression = myo_markers_expression %>% dplyr::filter(rownames(myo_markers_expression) != "IL12B") #remove IL12B since it has all 0
```

```{r results='hide'}
cor_mat = cor(expression %>% t(), myo_markers_expression %>% t())
cor_mat = cor_mat[complete.cases(cor_mat), ] #remove rows with NA in at least one col
# pheatmap(cor_mat[,1,drop = F],cluster_rows = T,cluster_cols = F,show_rownames = F, show_colnames = T,treeheight_row = 0)
top_10_TP63 =   cor_mat %>% as.data.frame %>% select(1) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
top_10_ACTA2 =   cor_mat %>% as.data.frame %>% select(2) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()
top_10_CNN1 =   cor_mat %>% as.data.frame %>% select(3) %>%  arrange(desc(1)) %>% top_n(10) %>% rownames()


all_top_genes = c(top_10_TP63,top_10_ACTA2,top_10_CNN1)
top_10_expression = expression %>% dplyr::filter(rownames(expression) %in% all_top_genes)
```

### Top correlated genes

```{r}
all_top_genes
```

### Heatmap 1

```{r}
colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(mat = cor(top_10_expression %>% t(), myo_markers_expression %>% t()),color = my_palette, breaks = colors)
```

### Heatmap 2

```{r}
colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(mat = cor(top_10_expression %>% t(), top_10_expression %>% t()),color = my_palette, breaks = colors)
```

### Enrichment analysis

```{r results='hide'}
genes_vec_enrichment(genes = all_top_genes,background = rownames(cor_mat),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "myo genes enrichment",add_msigdb_to_set = T)
```

##  {-}

## 1.c) genes that have >0.3 pearson with TP63, ACTA2 and CNN1 combined {.tabset}


```{r results='hide'}
cor_mat = cor(expression %>% t(), myo_markers_expression %>% t())
cor_mat = cor_mat[complete.cases(cor_mat), ] #remove rows with NA in at least one col

top_10_TP63 =    cor_mat %>% as.data.frame %>% select(1) %>% dplyr::filter(.>0.30) %>% rownames()
top_10_ACTA2 =   cor_mat %>% as.data.frame %>% select(2) %>% dplyr::filter(.>0.30) %>% rownames()
top_10_CNN1 =  cor_mat %>% as.data.frame %>% select(3) %>% dplyr::filter(.>0.30) %>% rownames()


all_top_genes = c(top_10_TP63,top_10_ACTA2,top_10_CNN1)
top_10_expression = expression %>% dplyr::filter(rownames(expression) %in% all_top_genes)
```

### top genes (first 30)
```{r}
all_top_genes %>% head(30)
```

### Heatmap 1

```{r}
colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(mat = cor(top_10_expression %>% t(), myo_markers_expression %>% t()),color = my_palette, breaks = colors)
```

### Heatmap 2

```{r}
colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(mat = cor(top_10_expression %>% t(), top_10_expression %>% t()),color = my_palette, breaks = colors)
```

### Enrichment analysis

```{r results='hide'}
genes_vec_enrichment(genes = all_top_genes,background = rownames(cor_mat),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "myo genes enrichment",add_msigdb_to_set = T)
```


##  {-}

# **Luminal markers**


```{r}
lum_protein_markers = c("KIT")
expression = GetAssayData(object = acc1_cancer_cells,assay = "RNA",slot = "data") %>% as.data.frame()
lum_markers_expression = expression %>% dplyr::filter(row.names(expression) %in% lum_protein_markers) %>% as.data.frame()
```

## 2.a) KIT expression
```{r}
FeaturePlot(object = acc1_cancer_cells,features = lum_protein_markers,min.cutoff = 0)
```


```{r results='hide'}
cor_mat = cor(expression %>% t(), lum_markers_expression %>% t())
cor_mat = cor_mat[complete.cases(cor_mat), ] #remove rows with NA in at least one col
# pheatmap(cor_mat[,1,drop = F],cluster_rows = T,cluster_cols = F,show_rownames = F, show_colnames = T,treeheight_row = 0)
top_10_KIT =   cor_mat %>% as.data.frame %>% select(1) %>%  arrange(desc(1)) %>% top_n(30) %>% rownames()
top_10_expression = expression %>% dplyr::filter(rownames(expression) %in% top_10_KIT)
```

## 2.b)top 30 correlated genes with KIT {.tabset}

### top genes
```{r}
top_10_KIT
```
### enrichment analysis
```{r results='hide'}
# hallmark_gene_set = msigdbr(species = "Homo sapiens")%>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()

genes_vec_enrichment(genes = top_10_KIT,background = rownames(expression),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "myo genes enrichment",add_msigdb_to_set = T,custom_pathways = luminal_gs)
```
## {-}

## 2.c)genes that have >0.3 pearson with KIT {.tabset}




```{r}
cor_mat = cor(expression %>% t(), lum_markers_expression %>% t())
cor_mat = cor_mat[complete.cases(cor_mat), ] #remove rows with NA in at least one col
top_10_KIT =   cor_mat %>% as.data.frame %>% select(1) %>% dplyr::filter(.>0.30) %>% rownames()
top_10_expression = expression %>% dplyr::filter(rownames(expression) %in% top_10_KIT)
```
### top genes
```{r}
top_10_KIT
```
### enrichment analysis

```{r results='hide'}
# hallmark_gene_set = msigdbr(species = "Homo sapiens")%>% dplyr::distinct(gs_name, gene_symbol) %>% as.data.frame()

genes_vec_enrichment(genes = top_10_KIT,background = rownames(expression),gene_sets = "homer_hallmark",convert_background= T,add_bg = F,title = "myo genes enrichment",add_msigdb_to_set = T,custom_pathways = luminal_gs)
```
## {-}



# **Option 2 - choosing from known markers**

## Results {.tabset}


### Lum genes expression

```{r fig.height=7, fig.width=11}
FeaturePlot(object = acc1_cancer_cells,features = lum_genes)
```
### Myo genes expression

```{r fig.height=7, fig.width=11}
FeaturePlot(object = acc1_cancer_cells,features = myo_genes)
```

### luminal score

```{r}
myoscore=apply(acc1_cancer_cells@assays[["RNA"]][c("TP63","TP73","CDH3","ACTA2","CDH11","TAGLN","MYLK","DKK3"),],2,mean)
lescore=apply(acc1_cancer_cells@assays[["RNA"]][c("KIT","CLDN3","ANXA8","ELF5","LCN2","WFDC2","ATP1B1"),],2,mean)
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,lescore-myoscore,"luminal_over_myo")
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,lescore,"luminal_score")
acc1_cancer_cells=AddMetaData(acc1_cancer_cells,myoscore,"myo_score")
```



```{r}
FeaturePlot(acc1_cancer_cells,features = "luminal_score",min.cutoff = 0)
```
### myo score
```{r}
FeaturePlot(acc1_cancer_cells,features = "myo_score",min.cutoff = 0)
```

### luminal_over_myo
```{r}
FeaturePlot(acc1_cancer_cells,features = "luminal_over_myo",min.cutoff = 0)
```
### luminal_over_myo2
```{r}
FeaturePlot(acc1_cancer_cells,features = "luminal_over_myo",cols = c("blue","grey","yellow"))
```
## {-}

# Option 3-  subset genes fron original ACC

## In HMSC {.tabset}

### Luminal genes:

```{r}
lum_markers  = FetchData(object = acc1_cancer_cells, vars = lum_genes)
lum_cor = cor(lum_markers)

colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(lum_cor,color = my_palette, breaks = colors,display_numbers = T)

```

### Myo genes:


```{r}
myo_markers  = FetchData(object = acc1_cancer_cells, vars = myo_genes)
myo_cor = cor(myo_markers)

colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(myo_cor,color = my_palette, breaks = colors,display_numbers = T)

```
## {-}

## In ACC {.tabset}

### Luminal genes:

```{r}
lum_markers  = FetchData(object = acc_cancer_cell_no_acc1, vars = lum_genes)
lum_cor = cor(lum_markers)

colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(lum_cor,color = my_palette, breaks = colors,display_numbers = T)

```

### Myo genes:


```{r}
myo_markers  = FetchData(object = acc_cancer_cell_no_acc1, vars = myo_genes)
myo_cor = cor(myo_markers)

colors <- c(seq(-1,1,by=0.01))
my_palette <- c("blue",colorRampPalette(colors = c("blue", "white", "red"))
                                                   (n = length(colors)-3), "red")
pheatmap(myo_cor,color = my_palette, breaks = colors,display_numbers = T)

```
## {-}
