---
title: '`r rstudioapi::getSourceEditorContext()$path %>% basename() %>% gsub(pattern = "\\.Rmd",replacement = "")`' 
author: "Avishai Wizel"
date: '`r Sys.time()`'
output:
  # html_notebook: 
  #   code_folding: hide
  #   toc: true
  #   toc_collapse: true
  #   toc_float:
  #     collapsed: false
  #   number_sections: true
  #   toc_depth: 1
  html_document: 
    code_folding: hide
    toc: yes
    toc_collapse: yes
    toc_float: 
      collapsed: TRUE
    number_sections: true
    toc_depth: 2
    df_print: paged

params:
  stratify: "M"
  log_tpm: TRUE
  data_out_dir: "./temp_data/" 
  figs_out_dir: NULL
  input_data: !r c(
    functions = "Notebooks/TCGA/TCGA_functions.R",
    opscc_deg =  "./Reports/HPV_OPSCC_Analysis_PMC10191634/02_findDEG/opscc_deg_5Kvargenes.rds",
    hmsc_deg = "./Reports/HMSC/05_HPV_analysis/HMSC_hpv_deg_df_noG2M_cells.csv",
    myb_tergets = "./Input_data/MYB_tergets_PMC4767593_sup_table_4.xlsx",
    tp53_data = "./Input_data/TCGA/cbioportal_hnsc_tcga_gdc_TP53_data.tsv",
    genesets_h = "./Input_data/h.all.v7.0.symbols.pluscc.gmt",
    HNSC_tpm = "./Reports/TCGA/build_TCGA_datasets/TCGA_HNSC_TPM.RDS",
    cbp_data = "./Input_data/TCGA/cbioportal_hnsc_tcga_pan_can_atlas_2018_clinical_data.tsv"
    global_params  = "./Input_data/global_params.R")
---

```{r}
# print params
paged_table(as.data.frame(unlist(params)))
```

# Data

```{r}
library(TCGAbiolinks)
library(biomaRt)
library(kableExtra)
library(rmarkdown)
library(readxl)
source(params$input_data["functions"])

```

```{r}
# load hpv signature genes
avg_log2FC_cutoff = log2(signature_fc)
fdr_cutoff = signature_fdr


opscc_deg = readRDS(file = params$input_data["opscc_deg"])
hmsc_deg = read.table(file = params$input_data["hmsc_deg"],row.names = 1,sep = ",",header = T)

hmsc_signatue = rownames(hmsc_deg[hmsc_deg$fdr<fdr_cutoff & hmsc_deg$avg_log2FC>avg_log2FC_cutoff,])


opscc_signature = opscc_deg %>% filter(avg_log2FC > avg_log2FC_cutoff &
                                  fdr < fdr_cutoff) %>% rownames()
opscc_signature

myb_tergets <- read_excel(params$input_data["myb_tergets"],col_names = T) %>%  pull(Gene)

tp53_data = read_tsv(file = params$input_data["tp53_data"],col_names = T)
  
```

```{r}
# load TPM data
HNSC_tpm = readRDS(params$input_data["HNSC_tpm"])

if (params$log_tpm) {
  HNSC_tpm = log2(HNSC_tpm+1)
}
```

```{r}
# load clinical data
project_name = "TCGA-HNSC"

clinical_data <- GDCquery_clinic(project_name, "clinical")
OPSCC_tissues = c("Base of tongue, NOS", #filter only OPSCC
"Oropharynx, NOS",
"Posterior wall of oropharynx",
"Tonsil, NOS")
clinical_data = clinical_data[clinical_data$tissue_or_organ_of_origin %in% OPSCC_tissues,]
```

```{r}
# add HPV statusfrom cbioportal to clinical_data
cbp_data = read_tsv(file = params$input_data["cbp_data"])
cbp_data = cbp_data[,c("Patient ID", "Subtype")] %>% dplyr::rename(hpv_status = Subtype) %>% 
  mutate(hpv_status = gsub(,x = hpv_status,pattern = "HNSC_HPV",replacement = "HPV"))
clinical_data = inner_join(x = clinical_data,y = cbp_data,by = c("submitter_id"="Patient ID"))
na_patients = clinical_data[is.na(clinical_data$hpv_status), "submitter_id"]
```

```{r}
# add TP53 mutation status from cbioportal to clinical_data

tp53_data = tp53_data[,c("Patient ID", "Altered")] %>% dplyr::rename(tp53_status = Altered) %>% 
  mutate(tp53_status = 
           case_when(tp53_status == 1 ~ "tp53_mutated",
                     tp53_status == 0 ~ "tp53_unmutated"))
clinical_data = inner_join(x = clinical_data,y = tp53_data,by = c("submitter_id"="Patient ID"))
na_patients = clinical_data[is.na(clinical_data$tp53_data), "submitter_id"]
```

# MYB

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method	=T
)

p[[1]]/p[[2]]
```

# HPV

```{r fig.height=6, fig.width=8}
clinical_data_for_hpv = clinical_data  %>% filter(!is.na(hpv_status))
p = TCGAanalyze_survival(
    data = clinical_data_for_hpv,
    clusterCol = "hpv_status",
    filename = NULL,pval.method = T
)

p[[1]]/p[[2]]
```

# MYB - HPV PPM correlation

```{r}
# load data 
library("readxl")
library(stringr)
library(ggpubr)
TCGA_HPV_data <- read_excel("./Input_data/TCGA/PMC3806554_ncomms3513-s5.xlsx",skip = 2)
TCGA_HPV_data$submitter_id = TCGA_HPV_data$`Sample Barcode` %>% str_sub( start = 1, end = 12) 
rownames(TCGA_HPV_data) = TCGA_HPV_data$submitter_id
TCGA_HPV_data = TCGA_HPV_data %>% filter(Cancer == "HNSC")
```

```{r}

gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status = gene_status[TCGA_HPV_data$submitter_id,] %>%  cbind(HPV_PPM = TCGA_HPV_data$ppm) %>% na.omit()

paged_table(gene_status)


sp <- ggscatter(gene_status, x = "MYB", y = "HPV_PPM",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
# Add correlation coefficient
sp + stat_cor(method = "pearson")

```

# HPV signature - HPV correlation

```{r}
# with HPV signature
gene_status = set_clinical_data(clin_data = clinical_data,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HPV+ signature")
gene_status = gene_status[TCGA_HPV_data$submitter_id,] %>% cbind(HPV_PPM = TCGA_HPV_data$ppm) %>% na.omit()
paged_table(gene_status)

sp <- ggscatter(gene_status, x = "HPV+ signature", y = "HPV_PPM",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
# Add correlation coefficient
sp + stat_cor(method = "pearson")


```

# HPV signature - HPV T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)
```

```{r}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(score ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "score", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") 

```

# MYB- HPV T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)



```

```{r fig.height=5, fig.width=7}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(MYB ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "MYB", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") +ylim(0,7)

```

# MYB and HPV survival plot

## HPV + tumors

```{r}
clinical_data_with_scores = clinical_data  %>% filter(hpv_status == "HPV+")

gene_status = set_clinical_data(clin_data = clinical_data_with_scores,genes = "NOTCH1",tpm_data_frame = HNSC_tpm,stratify = "T")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = inner_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))
# debugonce(TCGAanalyze_survival)

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```

## remove low smokers

```{r}
clinical_data_with_scores = clinical_data  %>% filter(hpv_status == "HPV+") %>% filter(pack_years_smoked < 100 | is.na(pack_years_smoked))

gene_status = set_clinical_data(clin_data = clinical_data_with_scores,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = "M")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = inner_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))
# debugonce(TCGAanalyze_survival)

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```

## HPV- tumors

```{r}
clinical_data_with_scores = clinical_data  %>% filter(hpv_status == "HPV-")

gene_status = set_clinical_data(clin_data = clinical_data_with_scores,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = inner_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```

## combine MYB and HPV survival plot

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)

clinical_data_with_scores = inner_join(x = clinical_data,y = gene_status, by = c("submitter_id"="patient"))

clinical_data_with_scores$gene_status = paste(clinical_data_with_scores$gene_status,clinical_data_with_scores$hpv_status,sep = ", ")
paged_table(clinical_data_with_scores)
```

```{r fig.height=6, fig.width=11}

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```

# Myb Targets

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = "NOTCH1",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)



```

```{r fig.height=5, fig.width=7}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(NOTCH1 ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "NOTCH1", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") +ylim(0,7)

```

# shared signature

```{r}
fileName =  
"./Rendered_notebooks/HPV_signature_compare/HPV_signature_compare_2024_09_08/shared_HPV_signature.txt"

shared_signature = scan(fileName, character(), quote = "")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = shared_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HPV+ signature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

p[[1]]/p[[2]]
```

## T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = shared_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)
```

```{r}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(score ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "score", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") 

```

# HMSC signature

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HPV+ signature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

p[[1]]/p[[2]]
```

```{r}
pdf(height=6,width=8,file = paste0(params$data_out_dir,"OPSCC_TCGA_HPV_signature_from_HMSC.pdf"))
p[[1]]/p[[2]]
dev.off()
```

## HPV +

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV+")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

(p[[1]]+ggtitle("OPSCC HPV+ samples"))/p[[2]]

```

## HPV -

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV-")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC HPV- samples"))/p[[2]]
```

## No cell cycle genes

```{r}
genesets_h  = geneIds(getGmt(params$input_data["genesets_h"]))
hmsc_signatue_no_cc = hmsc_signatue[!hmsc_signatue %in% genesets_h$HALLMARK_E2F_TARGETS]
```

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV-")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue_no_cc,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature no CC")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC HPV- samples"))/p[[2]]

```

## T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))
gene_status = gene_status %>% filter(!is.na(hpv_status))
paged_table(gene_status)



```

```{r}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(score ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "score", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") 

```

# OPSCC sc signature

## all samples

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "OPSCC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

(p[[1]]+ggtitle("OPSCC samples"))/p[[2]]

```

## HPV +

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV+")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "OPSCC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

(p[[1]]+ggtitle("OPSCC HPV+ samples"))/p[[2]]

```

## HPV -

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV-")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "OPSCC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC HPV- samples"))/p[[2]]
```

# Additional mutations

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = "TP53",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method	=T
)

p[[1]]/p[[2]]
```

## TP53

```{r fig.height=6, fig.width=8}
p = TCGAanalyze_survival(
    data = clinical_data,
    clusterCol = "tp53_status",
    filename = NULL,pval.method = T
)

p[[1]]/p[[2]]
```

## Fisher

```{r}
# plot
library(rstatix)

df = clinical_data[,c("tp53_status","hpv_status")]
test = fisher_test(table(df))
p = ggbarstats(
  df,
  tp53_status,
  hpv_status,
  results.subtitle = FALSE,
  subtitle = paste0("Fisher's exact test", ", p-value = ",
                    test$p)
)

p
```

## TP53 unmutated

```{r}
clinical_data_for_test = clinical_data  %>% filter(tp53_status == "tp53_unmutated")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_for_test,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_for_test[clinical_data_for_test$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC unmutated tp53 samples"))/p[[2]]
```

## TP53 mutated

```{r}
clinical_data_for_test = clinical_data  %>% filter(tp53_status == "tp53_mutated") 
```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_for_test,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature") 
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_for_test[clinical_data_for_test$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC mutated tp53 samples"))/p[[2]]
```
