---
title: 'OPSCC_TCGA' 
author: "Avishai Wizel"
date: '`r Sys.time()`'
output:
  # html_notebook:
  #   code_folding: hide
  #   toc: true
  #   toc_collapse: true
  #   toc_float:
  #     collapsed: false
  #   number_sections: true
  #   toc_depth: 1
  html_document:
    code_folding: hide
    toc: yes
    toc_collapse: yes
    toc_float:
      collapsed: TRUE
    number_sections: true
    toc_depth: 2
    df_print: paged

params:
  stratify: "M"
  log_tpm: TRUE
  days_to_show: 1825
  data_out_dir: "./temp_data/" 

---

```{r}
if (interactive()) {
  source("./pipe.R")
  input = get_input(get_current_path())
  output = get_output(get_current_path())
}
```

```{r}
paged_table(data.frame(value= unlist(params)))
```

# Data

```{r}
library(TCGAbiolinks)
library(biomaRt)
library(kableExtra)
library(rmarkdown)
library(readxl)
source(input$functions)
source(input$fc_params)
library(GSEABase)
library(grid)
library(ComplexHeatmap)
library(ggpubr)
library(stringr)
library(data.table)
library(ggstatsplot)
```

```{r}
# load hpv signature genes
avg_log2FC_cutoff = log2(signature_fc)
fdr_cutoff = signature_fdr


opscc_deg = read.table(file =  input$opscc_deg,row.names = 1,sep = ",",header = T)
hmsc_deg = read.table(file = input$hmsc_deg,row.names = 1,sep = ",",header = T)

hmsc_signatue = rownames(hmsc_deg[hmsc_deg$fdr<fdr_cutoff & hmsc_deg$avg_log2FC>avg_log2FC_cutoff,])


opscc_signature = opscc_deg %>% filter(avg_log2FC > avg_log2FC_cutoff &
                                  fdr < fdr_cutoff) %>% rownames()

message("opscc_signature:")
print(opscc_signature)

message("hmsc_signatue:")
print(hmsc_signatue)

myb_targets <- read_excel(input$myb_targets,col_names = T) %>%  pull(Gene)

tp53_data = read_tsv(file = input$tp53_data,col_names = T)
  
```

```{r}
# load TPM data
HNSC_tpm = readRDS(input$HNSC_tpm)

if (params$log_tpm) {
  HNSC_tpm = log2(HNSC_tpm+1)
}
```

```{r}
# load clinical data
project_name = "TCGA-HNSC"

clinical_data <- GDCquery_clinic(project_name, "clinical")
OPSCC_tissues = c("Base of tongue, NOS", #filter only OPSCC
"Oropharynx, NOS",
"Posterior wall of oropharynx",
"Tonsil, NOS")
clinical_data = clinical_data[clinical_data$tissue_or_organ_of_origin %in% OPSCC_tissues,]
```

```{r}
#add follow_up data
clinical_follow_up = read_tsv(input$clinical_follow_up) %>% as.data.frame()
#get only "Last Contact"
clinical_follow_up <- clinical_follow_up[clinical_follow_up$timepoint_category...135 == "Last Contact", ]

#add to clinical and rename days_to_last_follow_up> days_to_follow_up
clinical_data = inner_join(clinical_data,clinical_follow_up[,c("case_submitter_id","days_to_follow_up"),drop=F],by = c("submitter_id"="case_submitter_id")) %>% select(-days_to_last_follow_up) %>% dplyr::rename(days_to_last_follow_up = days_to_follow_up)

```

```{r}
# add HPV statusfrom cbioportal to clinical_data
cbp_data = read_tsv(file = input$cbp_data)
cbp_data = cbp_data[,c("Patient ID", "Subtype")] %>% dplyr::rename(hpv_status = Subtype) %>% 
  mutate(hpv_status = gsub(,x = hpv_status,pattern = "HNSC_HPV",replacement = "HPV"))
clinical_data = inner_join(x = clinical_data,y = cbp_data,by = c("submitter_id"="Patient ID"))
na_patients = clinical_data[is.na(clinical_data$hpv_status), "submitter_id"]
```

```{r}
# add TP53 mutation status from cbioportal to clinical_data

tp53_data = tp53_data[,c("Patient ID", "Altered")] %>% dplyr::rename(tp53_status = Altered) %>% 
  mutate(tp53_status = 
           case_when(tp53_status == 1 ~ "tp53_mutated",
                     tp53_status == 0 ~ "tp53_unmutated"))
clinical_data = inner_join(x = clinical_data,y = tp53_data,by = c("submitter_id"="Patient ID"))
na_patients = clinical_data[is.na(clinical_data$tp53_data), "submitter_id"]
```

# MYB

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method	=T, xlim = params$days_to_show
)

p[[1]]/p[[2]]
```
```{r}
pdf(paste0(params$data_out_dir,"KM_MYB_OPSCC.pdf"),width = 8,height = 6)
p[[1]]/p[[2]]
dev.off()
```

# HPV

```{r fig.height=6, fig.width=8}
clinical_data_for_hpv = clinical_data  %>% filter(!is.na(hpv_status))
p = TCGAanalyze_survival(
    data = clinical_data_for_hpv,
    clusterCol = "hpv_status",
    filename = NULL,pval.method = T
)

p[[1]]/p[[2]]
```

# MYB - HPV PPM correlation

```{r}
# load data 
library("readxl")
library(stringr)
library(ggpubr)
TCGA_HPV_data <- read_excel("./Input_data/TCGA/PMC3806554_ncomms3513-s5.xlsx",skip = 2)
TCGA_HPV_data$submitter_id = TCGA_HPV_data$`Sample Barcode` %>% str_sub( start = 1, end = 12) 
rownames(TCGA_HPV_data) = TCGA_HPV_data$submitter_id
TCGA_HPV_data = TCGA_HPV_data %>% filter(Cancer == "HNSC")
```

```{r}

gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status = gene_status[TCGA_HPV_data$submitter_id,] %>%  cbind(HPV_PPM = TCGA_HPV_data$ppm) %>% na.omit()

paged_table(gene_status)


sp <- ggscatter(gene_status, x = "MYB", y = "HPV_PPM",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
# Add correlation coefficient
sp + stat_cor(method = "pearson")

```

# HPV signature - HPV correlation

```{r}
# with HPV signature
gene_status = set_clinical_data(clin_data = clinical_data,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HPV+ signature")
gene_status = gene_status[TCGA_HPV_data$submitter_id,] %>% cbind(HPV_PPM = TCGA_HPV_data$ppm) %>% na.omit()
paged_table(gene_status)

sp <- ggscatter(gene_status, x = "HPV+ signature", y = "HPV_PPM",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   )
# Add correlation coefficient
sp + stat_cor(method = "pearson")


```

# HPV signature - HPV T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)
```

```{r}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(score ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "score", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") 

```

# MYB- HPV T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)



```


```{r fig.height=5, fig.width=7}
# plot
library(ggpubr)
library(rstatix)
gene_status$hpv_status = factor(gene_status$hpv_status,levels = c("HPV-","HPV+"))
stat.test <- gene_status %>%
  t_test(MYB ~ hpv_status) %>%
  add_significance()
stat.test

stat.test <- stat.test %>% add_xy_position(x = "hpv_status")


plt = ggplot(gene_status, aes(x = hpv_status, y = MYB)) +
  geom_violin(scale = "width",aes(fill = hpv_status)) +
  geom_boxplot(width = 0.2, outlier.shape = NA) +
  stat_pvalue_manual(stat.test, label = "{p.adj}")  + #add p value
  geom_jitter()+
  labs(title = "MYB expression by HPV status", y = "MYB expression: log2(TPM)", fill = "HPV Status")+
  theme(plot.title = element_text(hjust = 0.5),axis.title.x = element_blank())+scale_x_discrete(limits = c("HPV-","HPV+"))+theme_minimal()
plt
```


```{r}
pdf(paste0(params$data_out_dir,"violin_MYB_HPV_OPSCC_TCGA.pdf"),width = 7,height = 5)
plt
dev.off()
```
# MYB and HPV survival plot

## HPV + tumors

```{r}
clinical_data_with_scores = clinical_data  %>% filter(hpv_status == "HPV+")

gene_status = set_clinical_data(clin_data = clinical_data_with_scores,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = "T")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = inner_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))
# debugonce(TCGAanalyze_survival)

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```


## HPV- tumors

```{r}
clinical_data_with_scores = clinical_data  %>% filter(hpv_status == "HPV-")

gene_status = set_clinical_data(clin_data = clinical_data_with_scores,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = inner_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```

## combine MYB and HPV survival plot

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = "MYB",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)

clinical_data_with_scores = inner_join(x = clinical_data,y = gene_status, by = c("submitter_id"="patient"))

clinical_data_with_scores$gene_status = paste(clinical_data_with_scores$gene_status,clinical_data_with_scores$hpv_status,sep = ", ")
paged_table(clinical_data_with_scores)
```

```{r fig.height=6, fig.width=11}

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
p[[1]]/p[[2]]
```

# Myb Targets

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = "NOTCH1",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))

paged_table(gene_status)



```

```{r fig.height=5, fig.width=7}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(NOTCH1 ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "NOTCH1", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") +ylim(0,7)

```



# HMSC signature {.tabset}



```{r fig.height=6, fig.width=8,results='asis'}

# all samples:
gene_status = set_clinical_data(clin_data = clinical_data,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HPV+ signature")

clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

p1 = (p[[1]]+ggtitle("OPSCC samples"))/p[[2]]
print_tab(p1,title = "all samples")

pdf(paste0(params$data_out_dir,"all_samples_hmsc_signature.pdf"),width = 8,height = 6)
p1
dev.off()


# HPV+ samples:
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV+")
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature")
paged_table(gene_status)

clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

p1 = (p[[1]]+ggtitle("OPSCC HPV+ samples"))/p[[2]]
print_tab(p1,title = "HPV+")

pdf(paste0(params$data_out_dir,"HPVpos_samples_hmsc_signature.pdf"),width = 8,height = 6)
p1
dev.off()


# HPV neg samples:
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV-")
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature")
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

p1 = (p[[1]]+ggtitle("OPSCC HPV- samples"))/p[[2]]
print_tab(p1,title = "HPV-")

pdf(paste0(params$data_out_dir,"HPVneg_samples_hmsc_signature.pdf"),width = 8,height = 6)
p1
dev.off()
```






# {-}




## No cell cycle genes  {.tabset}

```{r fig.height=6, fig.width=8, results='asis'}

# remove genes:
genesets_h  = geneIds(getGmt(input$genesets_h))
message("cc genes in signature:")
print(hmsc_signatue[hmsc_signatue %in% genesets_h$HALLMARK_E2F_TARGETS])

hmsc_signatue_no_cc = hmsc_signatue[!hmsc_signatue %in% genesets_h$HALLMARK_E2F_TARGETS]

#all samples:

gene_status = set_clinical_data(clin_data = clinical_data,genes = hmsc_signatue_no_cc,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature no CC")

clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
p1 = (p[[1]]+ggtitle("OPSCC samples"))/p[[2]]
print_tab(p1,title = "OPSCC samples",subtitle_num = 3)

pdf(paste0(params$data_out_dir,"all_samples_hmsc_signature_noCC_genes.pdf"),width = 8,height = 6)
p1
dev.off()

#hpv +:
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV+")

gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue_no_cc,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature no CC")

clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
p1 = (p[[1]]+ggtitle("OPSCC HPV+ samples"))/p[[2]]
print_tab(p1,title = "HPV+",subtitle_num = 3)

pdf(paste0(params$data_out_dir,"HPVpos_samples_hmsc_signature_noCC_genes.pdf"),width = 8,height = 6)
p1
dev.off()

#hpv- :
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV-")

gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = hmsc_signatue_no_cc,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature no CC")

clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
p1 = (p[[1]]+ggtitle("OPSCC HPV- samples"))/p[[2]]
print_tab(p1,title = "HPV-",subtitle_num = 3)

pdf(paste0(params$data_out_dir,"HPVneg_samples_hmsc_signature_noCC_genes.pdf"),width = 8,height = 6)
p1
dev.off()
```



## T test

```{r}
# create data

gene_status = set_clinical_data(clin_data = clinical_data,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify)
gene_status$patient = rownames(gene_status)
gene_status = inner_join(x = gene_status,y = clinical_data[,c("submitter_id","hpv_status")],by = c("patient"="submitter_id"))
gene_status = gene_status %>% filter(!is.na(hpv_status))
paged_table(gene_status)



```

```{r}
# plot
library(ggpubr)
library(rstatix)
stat.test <- gene_status %>%
  t_test(score ~ hpv_status) %>%
  add_significance()
stat.test


bxp <- ggboxplot(gene_status, x = "hpv_status", y = "score", fill = "#00AFBB")+
  geom_jitter()
stat.test <- stat.test %>% add_xy_position(x = "hpv_status")
bxp + 
  stat_pvalue_manual(stat.test,  label = "T-test, p = {p}") 

```

# OPSCC sc signature

## all samples

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "OPSCC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

(p[[1]]+ggtitle("OPSCC samples"))/p[[2]]

```

## HPV +

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV+")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "OPSCC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set",
    height = 10,
    width=10,filename = NULL,pval.method = T
)

(p[[1]]+ggtitle("OPSCC HPV+ samples"))/p[[2]]

```

## HPV -

```{r}
clinical_data_hpv_pos = clinical_data  %>% filter(hpv_status == "HPV-")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_hpv_pos,genes = opscc_signature,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "OPSCC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_hpv_pos[clinical_data_hpv_pos$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC HPV- samples"))/p[[2]]
```

# Additional mutations

```{r}
gene_status = set_clinical_data(clin_data = clinical_data,genes = "TP53",tpm_data_frame = HNSC_tpm,stratify = params$stratify)
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data[clinical_data$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    main = "TCGA Set\n GBM",
    height = 10,
    width=10,filename = NULL,pval.method	=T, xlim = params$days_to_show
)

p[[1]]/p[[2]]
```

## TP53

```{r fig.height=6, fig.width=8}
p = TCGAanalyze_survival(
    data = clinical_data,
    clusterCol = "tp53_status",
    filename = NULL,pval.method = T
)

p[[1]]/p[[2]]
```

## Fisher

```{r}
# plot
library(rstatix)

df = clinical_data[,c("tp53_status","hpv_status")]
test = fisher_test(table(df))
p = ggbarstats(
  df,
  tp53_status,
  hpv_status,
  results.subtitle = FALSE,
  subtitle = paste0("Fisher's exact test", ", p-value = ",
                    test$p)
)

p
```

## TP53 unmutated

```{r}
clinical_data_for_test = clinical_data  %>% filter(tp53_status == "tp53_unmutated")

```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_for_test,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature")
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_for_test[clinical_data_for_test$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC unmutated tp53 samples"))/p[[2]]
```

## TP53 mutated

```{r}
clinical_data_for_test = clinical_data  %>% filter(tp53_status == "tp53_mutated") 
```

```{r}
gene_status = set_clinical_data(clin_data = clinical_data_for_test,genes = hmsc_signatue,tpm_data_frame = HNSC_tpm,stratify = params$stratify,signature_name = "HMSC HPV \nsignature") 
paged_table(gene_status)
```

```{r fig.height=6, fig.width=8}
clinical_data_with_scores = clinical_data_for_test[clinical_data_for_test$submitter_id %in% rownames(gene_status),]
gene_status$patient = rownames(gene_status)
clinical_data_with_scores = left_join(x = clinical_data_with_scores,y = gene_status, by = c("submitter_id"="patient"))

p = TCGAanalyze_survival(
    data = clinical_data_with_scores,
    clusterCol = "gene_status",
    height = 10,
    width=10,filename = NULL,pval.method = T
)
(p[[1]]+ggtitle("OPSCC mutated tp53 samples"))/p[[2]]
```


# session info
<button class="btn btn-primary" data-toggle="collapse" data-target="#BlockName"> Show/Hide </button>  
<div id="BlockName" class="collapse">  
```{r}
library(devtools)
session_info()
```

</div>